\normalfont \normalsize \textsc{University of Stuttgart} \\ [25pt] \horrule{0.5pt} \\[0.4cm] \huge Hauptseminar:\\The Logjam Attack \\ \horrule{2pt} \\[0.5cm]

### 2017-06-21 Ausarbeitung/paper.tex
final version v1.0

\title{	
\normalfont \normalsize 
\textsc{University of Stuttgart} \\ [25pt] % Your university, school and/or department name(s)
\horrule{0.5pt} \\[0.4cm] % Thin top horizontal rule
\huge Hauptseminar:\\The Logjam Attack \\ % The assignment title
\horrule{2pt} \\[0.5cm] % Thick bottom horizontal rule
}

@@

\begin{document}

\maketitle % Print the title

\tableofcontents

\newpage

%----------------------------------------------------------------------------------------
%	ABSTRACT
%----------------------------------------------------------------------------------------
\section*{Abstract}
\begin{abstract}
In the field of study of security for computer networks, \citep{Adrian:2015:IFS:2810103.2813707} developed a new attack called Logjam to break many cryptographic systems that use Diffie-Hellman key exchange mechanism, where TLS 1.2 was the current standard at that time to establish secure end-to-end connections. This work elucidates technical details of the background for Logjam, in particular Diffie-Hellman key exchange, Number Field Sieve algorithm and TLS handshake, and summarizes the overall results.
\end{abstract}
%----------------------------------------------------------------------------------------
%	INTRODUCTION
%----------------------------------------------------------------------------------------
\section{Introduction}
If it comes to network security, the Diffie-Hellman key exchange is the standard wildly used in practice and declared to be safe. However, \citep{Adrian:2015:IFS:2810103.2813707} could break the security on many popular and browser certified websites. They demonstrated this so called Logjam attack on the Transfer Level Security protocol that supports Diffe-Hellman as key exchange. They also implemented an attack exploiting weak and misconfigured groups using the Pohling-Hellman algorithm.

The second result the authors present is an estimate on cost of computing logs from key sizes of 1024 bits. I will briefly summarize their assumptions for the estimate and the results.
%----------------------------------------------------------------------------------------
%	Diffie-Hellman Key Exchange
%----------------------------------------------------------------------------------------
\section{Diffie-Hellman Key Exchange}
\label{sec:DHKeyExchange}
\citep{diffie1976new} was the first paper presenting a practical solution to the privacy and the authentication problem that does not rely on secure communication channels. These kinds of channels are often not given or so inefficient to use that the benefit of telecommunication is nullified. Hence, it was a historical and great advance in computer networking and cryptography. They presented an asymmetric approach with trap door functions that have an exponential cipher-cryptanalyst ratio. Because of this, a security system can choose feasible large private keys that are at the same time infeasible to infer from public keys and encrypted messages w.r.t. computation time.
\subsection{Problem Formulation}
Given:
\begin{itemize}
	\item $A, B$, the communication partners.
	\item Only insecure communication channels are available.
	\item $\{M\}$, a finite message space which contain all possible messages $M$ to be exchanged.
\end{itemize}
Determine:
\begin{itemize}
	\item $C = (\{E_K\}_{K\in\{K\}}, \{D_K\}_{K\in\{K\}})$, a \textit{public key cryptosystem} with mutual inverse functions
	\begin{itemize}
		\item $E_K:\{M\}\rightarrow\{M\}$
		\item $D_K:\{M\}\rightarrow\{M\}$
	\end{itemize}
	denoted as \textit{encryption} and \textit{decryption transformation} respectively.
	\item $\{K\}$, a suitable set of keys.
	\item Such that
	\begin{enumerate}
	\item $\forall K\in\{K\}:M = D_K(E_K(M))$
	\item $\forall K\in\{K\},M\in\{M\}:E_K(M), D_K(M)$ are easy to compute.
	\item $\forall^\infty K\in\{K\},\alpha:\{E_K\}_{K\in\{K\}}\rightarrow \{D_K\}_{K\in\{K\}}:D_K = \alpha(E_K)$ is infeasible to compute.
	\item $\forall K\in\{K\}\exists\beta:\{K\}\rightarrow(\{E_K\}_{K\in\{K\}}, \{D_K\}_{K\in\{K\}}):\beta(K) = (E_K,D_K)$ is feasible to compute.
	\end{enumerate}
\end{itemize}

\subsection{Solutions}
\paragraph{Key Exchange}
The solution the authors propose is to draw private and public keys $(D,E)$ from finite fields $\mathrm{GF}(q)\cong\mathbb{Z}/q\mathbb{Z}$, i.e.
\begin{equation}
D, E \in \mathrm{GF}(q) = \{x \mod q \mid x \in \mathbb{Z}\}.
\end{equation}
By choosing the private key as the logarithm of the public key and an arbitrary basis $\alpha \in \mathrm{GF}(q)$,
\begin{eqnarray}
D = \alpha^E\mod q,\\
E = \log_\alpha Y \mod q,\\
1 \leq D,E \leq q-1,
\end{eqnarray}
the requirements 3. and 4. are fulfilled. Calculating $D$ from $E$ has logarithmic time complexity in $q$. Inversely, calculating $E$ from $D$ has a complexity of $\sqrt{q}$. Therefore the cipher-cryptanalyst ratio is exponential. Note that this is not a proven bound. We use the best known algorithms for these problem to get this ratio. Better algorithms have not been found for decades, therefore this method is considered safe. Also, one need to find good primes $q$ in order to assure worst or near worst case log-computation time.

Now if $A$ and $B$ want to exchange a session key, they first choose public and private keys $D_A$, $D_B$, $E_A$, $E_B$ from a number field $\mathrm{GF}(q)$ with a sufficient size $q$ and send each other their public keys over the insecure communication channel. Then, they can calculate the common session key
\begin{eqnarray}
K_{A,B} & = & \alpha^{E_AE_B} \mod q\\
& = & (\alpha^{E_A})^{E_B} \mod q\\
& = & {D_A}^{E_B} \mod q\label{eqn:keyA}\\
& = & (\alpha^{E_B})^{E_A} \mod q\\
& = & {D_B}^{E_A} \mod q.\label{eqn:keyB}
\end{eqnarray}
While $A$ calculates the right term in \ref{eqn:keyA}, $B$ calculates the right term in \ref{eqn:keyB} and both of them have now the same session key $K_{A,B}$.

Because of the trap-door property of this mechanism, it allows us to use it for one-way authentication. A signature $s$ of some user $A$ is $E_A(s)$. No one can forge it, since $E_A$ is private. $A$ cannot deny his or her signature, since everyone can confirm $D_A(E_A(s)) = s$.

\paragraph{Cryptanalysis}
The attacker can compute $K_{A,B}$ either by computing $E_A$ from $D_B$ or $E_B$ from $D_A$, both are log-problems in finite fields. Thus, the attacker is successful, iff he or she can effort exponentially more computation power that $A$ and $B$.

%----------------------------------------------------------------------------------------
%	Number Field Sieve Algorithm
%----------------------------------------------------------------------------------------
\section{Number Field Sieve Algorithm}
As described above, the only known way to break Diffie-Hellman is to compute $E$ from $D$, i.e. to compute discrete logs for prime fields for a fixed prime $q$ and base $\alpha$. \citep{gordon1993discrete} first proposed how the \textit{number field sieve} (NFS) algorithm can be used to compute discrete logs. Before that, this technique was known to solve the factorization problem for large primes.

\subsection{Approach}
The idea is to factorize $D$ and find logs for each of these factors. The logs of these factors can be constructed by the logs of some precomputed logs in the database. This step can be done very efficiently. The difficult problem is to build up such a database for the prime of interest $q$. But the good thing is that this precomputation only depends on $q$ and not on a specific problem instance $D$. Once the precomputation is done, the database can be used to compute the log for every $D\in \mathrm{GF}(q)$.

@@

\subsubsection{Smoothness Assumption}
Define that an integer number $x$ is $B$-smooth, if all prime factors of $x$ are smaller that $B$. NFS assumes the base $\alpha$ to be $B$-smooth where $B$ is an upper bound for the prime values in the factor base (which is the output of the precomputation). If $\alpha$ is not $B$-smooth, we can use a $B$-smooth $\alpha'$ instead. Then, we can compute $\log_{\alpha'}\alpha$ and transform back to the original problem with
\begin{equation}
\log_\alpha D\equiv\frac{\log_{\alpha'}D}{\log_{\alpha'}\alpha}\mod (q-1)
\text{ .}
\end{equation}
If $\alpha'$ is not a generator for $\mathrm{GF}(q)^*$, the logs on the right hand side might not exist. $\alpha'$ can be checked for feasibility by factoring $(q-1)$ with the NFS factoring algorithm and check
\begin{equation}
(a')^{\frac{q-1}{p}} \not\equiv 1 \mod q,\quad\forall p\parallel q
\text{ ,}
\end{equation}
where $p\parallel q$ denotes that $p$ is a prime divisor of $q$. The reason why $\alpha'$ needs to be $B$-Smooth is that the factor base needs to have at least one inhomogeneous relation for the logs of the factor base, using the equation
\begin{eqnarray}
\label{eqn:assumption}
\log_\alpha\alpha = 1 \equiv \sum_{p^t\parallel a}t\log_\alpha p \mod (q-1)
\text{ ,}
\end{eqnarray}
otherwise we gain no knowledge from the factor base.

@@

Next, let $\gamma\in\mathbb{C}$ be the root of $f$, define $K = \mathbb{Q}(\gamma) = \mathbb{Q}\cup\{\gamma\}$ and $O_K$ the ring of integers in $K$. Since $(q,\Delta_f) = 1$, $\hat{q}:=(q,\gamma-m)$ is a first-degree prime factor of $q\in O_K$. It follows that $O_K/\hat{q}\cong\mathrm{GF}(q)$ and we can define a homomorphism $\phi:\mathbb{Z}[\alpha]\rightarrow\mathbb{Z}/q\mathbb{Z}$, providing us the representation we aimed for. We denote a prime ideal $s\in O_K$ as \textit{good} if $s$ does not divide the index $[O_K:\mathbb{Z}[\gamma]]$, otherwise \textit{bad}.

The actual factor base $\mathcal{B}$ for which we want to find the logarithms consists of two parts: $\mathcal{B}_\mathbb{Q}$ the set of rational primes and $\mathcal{B}_K$ the set of good prime ideals in $O_K$, where all of these elements are bounded in size, i.e. $p,\Vert p'\Vert_1\leq B,\ \forall p\in\mathcal{B}_\mathbb{Q},p'\in\mathcal{B}_K$. Note that $B$ is the same bound for our base $\alpha$ (from our primary problem definition: Find $E$ for $D=\alpha^E\mod q$). Also, let $\mathcal{B}'\subset\mathcal{B}_\mathbb{Q}$ denote the set of prime factors of $\alpha$.

@@

From here we can derive factorizations for primes in $\mathcal{B}_K$ with the fact (not proven here) that there is an unique ideal $\hat{s}\in\mathcal{B}_K$ lying over $s$ and dividing $c+d\gamma$. Let $v_{\hat{S}}(c,d) = v_s(c,d)$ for this ideal and be zero for other ideals in $\mathcal{B}_K$ of norm $s$. Then we have the following
\begin{eqnarray}
\label{eqnSieve1}
c+dm = \prod_{s\in\mathcal{B}_\mathbb{Q}}s^{w_s(c,d)}
\end{eqnarray}
and
\begin{eqnarray}
\label{eqnSieve2}
c+d\gamma = \prod_{\hat{s}\in\mathcal{B}_K}\hat{s}^{v_{\hat{s}}(c,d)}
\text{ .}
\end{eqnarray}
Repeat the steps described so far to find more pairs $(c,d)$ to build equations \ref{eqnSieve1} and \ref{eqnSieve2} until we have at least $|\mathcal{B}|$ of these equations. For each equation, build a matrix, where its rows are the $w_s$'s and $v_{\hat{s}}$'s, we call this set $\mathcal{A}$. We modify $\mathcal{A}$ by extracting only those columns where elements are in $\mathcal{B}-\mathcal{B}'$. This means, we are only interested in solutions for primes $s\in\mathcal{B}'$. $\mathcal{A}$ is now the input for the next step.

\subsubsection{Linear Algebra}
For each $A \in \mathcal{A},\ A = \mathbb{Z}^{S\times T},\ S > T$, where each row has at most $E$ non-zero entries, find a linear relation over $\mathbb{Q}$ for the rows of $A$. This is accomplished in four steps, shown for the case $S=T+1$. This procedure was developed by Carl Pomerance (*1944) which was not separately published, instead only for helping out in \citep{gordon1993discrete}.
\paragraph{Step 1}Attempt to compute the rank $r = \mathrm{rk}(A)$. For that, choose a random prime $q_0\leq ET\log T$. Later, we will check if that choice was feasible. If not, come back here and sample another one. Then, use Gaussian elimination mod $q_0$ to find the rank $r_0$ of $A\mod q_0$. Rearrange the rows so that the first $r_0$ rows are linearly independent mod $q_0$. Now the rearranged matrix is
\begin{equation}
A' =
\begin{pmatrix}
v_1\\v_2\\\vdots\\v_{T\!+\!1}
\end{pmatrix}
=
\begin{pmatrix}
a'_{11} & a'_{12} & \cdots & a'_{1T} \\
a'_{21} & a'_{22} & \cdots & a'_{2T} \\
\vdots & \vdots & \ddots & \vdots \\
a'_{T\!+\!11} & a'_{T\!+\!12} & \cdots & a'_{T\!+\!1T} \\
\end{pmatrix}
\text{ .}
\end{equation}
The result of the Gaussian elimination provide a submatrix $\hat{A} = \mathbb{Z}^{r_0\times r_0}$ of the first $r_0$ rows and columns that is nonsingular mod $q_0$,
\begin{equation}
\hat{A} =
\begin{pmatrix}
v_1\\v_2\\\vdots\\v_{r_0}
\end{pmatrix}
=
\begin{pmatrix}
a'_{11} & a'_{12} & \cdots & a'_{1r_0} \\
a'_{21} & a'_{22} & \cdots & a'_{2r_0} \\
\vdots & \vdots & \ddots & \vdots \\
a'_{r_01} & a'_{r_02} & \cdots & a'_{r_0r_0} \\
\end{pmatrix}
\text{ .}
\end{equation}
\paragraph{Step 2} Attempt to express $v_{r_0+1}$ as a linear combination of $v_1,...,v_{r_0}\mod q$ for each $q\leq ET\log T$. For some $q$ it will be successful, for others not. Denote the product of all successful primes as $\boldsymbol{P}$ and the product of all unsuccessful primes as $\boldsymbol{P'}$. If $\boldsymbol{P'} > ET\log T$, then repeat Step 1 (sample a new random prime number). Otherwise, continue with the next step.
\paragraph{Step 3} Attempt to find the determinant $D$ of $\hat{A}$. For each prime $q|\boldsymbol{P}$, use the Wieldemann's probabilistic determinant algorithm \citep{wiedemann1986solving} to compute an integer $D_q\in\{0,...,q-1\}$, which is the determinant of $A\mod q$ with probability of at least $1-(ET)^{-2}$, i.e.
\begin{equation}
\mathbb{P}(D_q = \det(A\mod q)\mid A,q) \geq 1-(ET)^{-2}
\text{ .}
\end{equation}
For each $D_q$ we can write an congruence
\begin{equation}
D_0 \equiv D_q \mod q,\quad\forall q|\boldsymbol{P}
\end{equation}
and use the Chinese remainder theorem to compute $D_0$ fulfilling these congruences and being closest to zero. Repeat this step until we have found a value that lies within the bound
\begin{equation}
0 < |D_0| \leq (E^{\frac{1}{2}}T)^T
\text{ .}
\end{equation}

\paragraph{Step 4} In the final step we try to find the actual coefficients $c_1,...,c_{r_0}$ for the linear relation between $v_{r_0+1}$ and the rest of the rows. Again, use the Chinese remainder theorem to find these coefficients closest to zero such that
\begin{equation}
D_0v_{r_0+1}\equiv\sum_{i=1}^{r_0}c_iv_i\mod\boldsymbol{P}
\text{ .}
\end{equation}
In case one of these coefficients $c_i$ exceeds $(E^{\frac{1}{2}}T)^T$ in absolute value, repeat Step 3 to find another $D_0$. Otherwise, we have found the relation
\begin{equation}
D_0v_{r_0+1}=\sum_{i=1}^{r_0}c_iv_i
\text{ .}
\end{equation}
It can also be shown that the expected runtime of this algorithm is
\begin{equation}
\mathcal{O}(E^2T^3\log^3T)
\text{ .}
\end{equation}

@@

After applying the algorithm described in the Linear Algebra step on the matrices resulted form the Sieving step, we have some new equations
\begin{equation}
\prod_{(c,d)}(c+dm)^{x(c,d)} = U
\textit{ ,}
\end{equation}
where $x(c,d)$ is the solution for the matrix constructed from $(c,d)$ and $U$ is a unit in $O_K$. After some intermediate steps, we can express the logs of primes $s\in\mathcal{B}$ as $z_s$ (which is a black box here) and state
\begin{equation}
\prod_{s\in\mathcal{B}'}s^{z_s}\equiv 1\mod q
\text{ .}
\end{equation}
Taking the log, we have
\begin{equation}
\sum_{s\in\mathcal{B}'}z_s\log_\alpha s \equiv 0 \mod (q-1)
\text{ .}
\end{equation}
With more than $|\mathcal{B}'|$ such congruences, we can solve these with equation \ref{eqn:assumption} and obtain the logs for all primes $s\in\mathcal{B}'$.

>>>>

\title{	
\normalfont \normalsize 
\textsc{University of Stuttgart} \\ [25pt] % Your university, school and/or department name(s)
\horrule{0.5pt} \\[0.4cm] % Thin top horizontal rule
\huge Hauptseminar: Technical Details on the Logjam Attack \\ % The assignment title
\horrule{2pt} \\[0.5cm] % Thick bottom horizontal rule
}

@@

\begin{document}
\pagenumbering{gobble}% Remove page numbers (and reset to 1)
\clearpage
\thispagestyle{empty}
\maketitle % Print the title

\newpage
\tableofcontents

\newpage
\clearpage
\pagenumbering{arabic}% Arabic page numbers (and reset to 1)
%----------------------------------------------------------------------------------------
%	ABSTRACT
%----------------------------------------------------------------------------------------
\section*{Abstract}
\begin{abstract}
In the field of study of security for computer networks, \citep{Adrian:2015:IFS:2810103.2813707} developed a new attack called Logjam to break many cryptographic systems that use Diffie-Hellman key exchange, where TLS 1.2 was the current standard at that time to establish secure end-to-end connections. This work elucidates technical details on the background of Logjam, in particular Diffie-Hellman key exchange, Number Field Sieve algorithm and TLS handshake, and summarizes the overall results.
\end{abstract}
%----------------------------------------------------------------------------------------
%	INTRODUCTION
%----------------------------------------------------------------------------------------
\section{Introduction}
If it comes to network security, Diffie-Hellman key exchange is the standard wildly used in practice and declared to be secure. However, \citep{Adrian:2015:IFS:2810103.2813707} could break the security on many popular and browser certified websites. They demonstrated this so called Logjam attack on the Transfer Level Security protocol that supports Diffe-Hellman as key exchange. They also implemented an attack exploiting the use of small exponents with unsafe primes using a combination of the Pohling-Hellman decomposition and Pollard's lambda method.

Section 2 provide a formal definition of the key exchange problem using asymmetric keys and describes the solution according to \citep{diffie1976new}. In order to break Diffie-Hellman, \citep{gordon1993discrete} first proposed how the Number Field Sieve algorithm can be used to compute logarithms in prime fields. This method is rather complicated and covers a large part of this work in Section 3. While most methods presented here is about cryptography, Section 4 presents the Transport Layer Security, specified in \citep{dierks2008transport}, which is a security protocol used in the transport layer from the OSI model. This model is the most commonly used one that provide a decomposition of a network system into layers, from physical to application level. Section 5 uses the methods from the previous sections to explain Logjam in short. Section 6 presents an attack proposed by \citep{van1996diffie} that exploits a specific kind of bad parameter choices made by system designers when Diffie-Hellman key exchange is used. The second result presented in \citep{Adrian:2015:IFS:2810103.2813707} is an estimate on the cost of computing logs of group elements using 1024 bits primes. I will briefly summarize their assumptions for the estimate and their results in Section 7. A conclusion is drawn about what is necessary to make computer networks more secure in the future in Section 8.
%----------------------------------------------------------------------------------------
%	Diffie-Hellman Key Exchange
%----------------------------------------------------------------------------------------
\section{Diffie-Hellman Key Exchange}
\label{sec:DHKeyExchange}
\citep{diffie1976new} was the first paper presenting a practical solution to the privacy and the authentication problem that does not rely on secure communication channels. These kinds of channels are often not given or so inefficient to use that the benefit of telecommunication is nullified. Hence, it was a historical and great advance in computer networking and cryptography. They presented an asymmetric approach with trap door functions that have an exponential cipher-cryptanalyst ratio. Because of this, a security system can choose feasible large private keys that are at the same time infeasible to infer from public keys and encrypted messages w.r.t. computation time.
\subsection{Problem Formulation}
Given:
\begin{itemize}
	\item $A, B$, the communication partners.
	\item Only insecure communication channels are available.
	\item $\{M\}$, a finite message space which contain all possible messages $M$ to be exchanged.
\end{itemize}
Determine:
\begin{itemize}
	\item $C = (\{E_K\}_{K\in\{K\}}, \{D_K\}_{K\in\{K\}})$, a \textit{public key cryptosystem} with mutual inverse functions
	\begin{itemize}
		\item $E_K:\{M\}\rightarrow\{M\}$
		\item $D_K:\{M\}\rightarrow\{M\}$
	\end{itemize}
	denoted as \textit{encryption} and \textit{decryption transformation} respectively.
	\item $\{K\}$, a suitable set of keys.
	\item Such that
	\begin{enumerate}
	\item $\forall K\in\{K\}:M = D_K(E_K(M))$
	\item $\forall K\in\{K\},M\in\{M\}:E_K(M), D_K(M)$ are easy to compute.
	\item $\forall^\infty K\in\{K\},f:\{E_K\}_{K\in\{K\}}\rightarrow \{D_K\}_{K\in\{K\}}:D_K = f(E_K)$ is infeasible to compute.
	\item $\forall K\in\{K\}\exists g:\{K\}\rightarrow(\{E_K\}_{K\in\{K\}}, \{D_K\}_{K\in\{K\}}):g(K) = (E_K,D_K)$ is feasible to compute.
	\end{enumerate}
\end{itemize}
The goal is therefore to find a key space $\{K\}$ and one transformation generator $g$ to generate public and private keys $(E_K,D_K)$ that fulfil these four conditions above.

The reason we call the public and private keys "encryption" and "decryption" transformations is that these keys can be used for secure message exchange directly. Everyone who wants to send a message $M$ to, say, $A$ just needs to send $E_A(M)$ and $A$ can read $D_A(M)$. This however is slower than cipher algorithms using symmetric keys. For the sake of performance, the solution proposed uses this asymmetric mechanism to exchange a session key, that is then used to encrypt application data.

\subsection{Solutions}
\paragraph{Key Exchange}
The solution the authors propose is to draw public and private keys $(E,D)$ from finite fields $\mathrm{GF}(q)\cong\mathbb{Z}/q\mathbb{Z}$, i.e.
\begin{equation}
E, D \in \mathrm{GF}(q) = \{x \mod q \mid x \in \mathbb{Z}\}.
\end{equation}
By choosing the private key as the logarithm of the public key and an arbitrary basis $\alpha \in \mathrm{GF}(q)$,
\begin{eqnarray}
E = \alpha^D\mod q,\\
D = \log_\alpha E \mod q,\\
1 \leq D,E \leq q-1,
\end{eqnarray}
the requirements 3. and 4. are fulfilled. Calculating $E$ from $D$ has logarithmic time complexity in $q$. Inversely, calculating $D$ from $E$ has a complexity of $\sqrt{q}$. Therefore the cipher-cryptanalyst ratio is exponential. Note that this is not a proven bound. We use the best known algorithms for these problem to get this ratio. Better algorithms have not been found for decades, therefore this method is considered safe. Also, one need to find good primes $q$ in order to assure worst or near worst case log-computation time.

Now if $A$ and $B$ want to exchange a session key, they first choose public and private keys $E_A$, $E_B$, $D_A$, $D_B$ from a number field $\mathrm{GF}(q)$ with a sufficient size $q$ and send each other their public keys over the insecure communication channel. Then, they can calculate the common session key
\begin{eqnarray}
K_{A,B} & = & \alpha^{D_AD_B} \mod q\\
& = & (\alpha^{D_A})^{D_B} \mod q\\
& = & {E_A}^{D_B} \mod q\label{eqn:keyA}\\
& = & (\alpha^{D_B})^{D_A} \mod q\\
& = & {E_B}^{D_A} \mod q.\label{eqn:keyB}
\end{eqnarray}
While $A$ calculates the right term in \ref{eqn:keyA}, $B$ calculates the right term in \ref{eqn:keyB} and both of them have now the same session key $K_{A,B}$.

Because of the trap-door property of this mechanism, it allows us to use it for one-way authentication. A signature $s$ of some user $A$ is $D_A(s)$. No one can forge it, since $D_A$ is private. $A$ cannot deny his or her signature, since everyone can confirm $E_A(D_A(s)) = s$.

\paragraph{Cryptanalysis}
The attacker can compute $K_{A,B}$ either by computing $D_A$ from $E_B$ or $D_B$ from $E_A$, both are log-problems in finite fields. Thus, the attacker is successful iff he or she can effort exponentially more computation power that $A$ and $B$.

%----------------------------------------------------------------------------------------
%	Number Field Sieve Algorithm
%----------------------------------------------------------------------------------------
\section{Number Field Sieve Algorithm}
As described above, the only known way to break Diffie-Hellman is to compute $D$ from $E$, i.e. to compute discrete logs for prime fields for a fixed prime $q$ and base $\alpha$. \citep{gordon1993discrete} first proposed how the \textit{Number Field Sieve} (NFS) algorithm can be used to compute discrete logs. Before that, this technique was known to solve the factorization problem for large primes.

\subsection{Approach}
The idea is to factorize $E$ and find logs for each of these factors. The logs of these factors can be constructed by the logs of some precomputed logs in the database. This step can be done very efficiently. The difficult problem is to build up such a database for the prime of interest $q$. But the good thing is that this precomputation only depends on $q$ and not on a specific problem instance $E$. Once the precomputation is done, the database can be used to compute the log for every $E\in \mathrm{GF}(q)$.

@@

\subsubsection{Smoothness Assumption}
Define that an integer number $x$ is $B$-smooth, if all prime factors of $x$ are smaller than $B$. NFS assumes the base $\alpha$ to be $B$-smooth where $B$ is an upper bound for the prime values in the factor base (which is the output of the precomputation). $B$ is restricted by our computation capabilities. If $\alpha$ is not $B$-smooth, we can use a $B$-smooth $\alpha'$ instead. Then, we can compute $\log_{\alpha'}\alpha$ and transform back to the original problem with
\begin{equation}
\log_\alpha E\equiv\frac{\log_{\alpha'}E}{\log_{\alpha'}\alpha}\mod (q-1)
\text{ .}
\end{equation}
If $\alpha'$ is not a generator for $\mathrm{GF}(q)^*$, the logs on the right hand side might not exist. $\alpha'$ can be checked for feasibility by factoring $(q-1)$ with the NFS factoring algorithm and check
\begin{equation}
(a')^{\frac{q-1}{p}} \not\equiv 1 \mod q,\quad\forall p\parallel q
\text{ ,}
\end{equation}
where $p\parallel q$ denotes that $p$ is a prime divisor of $q$. The reason why $\alpha$ or $\alpha'$ needs to be $B$-Smooth is that the factor base needs to have at least one inhomogeneous relation for the logs of the factor base, using the equation
\begin{eqnarray}
\label{eqn:assumption}
\log_\alpha\alpha = 1 \equiv \sum_{p^t\parallel a}t\log_\alpha p \mod (q-1)
\text{ ,}
\end{eqnarray}
which is not guaranteed to be the case if not all log factors are considered.

@@

Next, let $\gamma\in\mathbb{C}$ be the root of $f$, define $K = \mathbb{Q}(\gamma) = \mathbb{Q}\cup\{\gamma\}$ and $O_K$ the ring of integers in $K$. Since $\mathrm{gcd}(q,\Delta_f) = 1$, $\hat{q}:=(q,\gamma-m)$ is a first-degree prime factor of $q\in O_K$. It follows that $O_K/\hat{q}\cong\mathrm{GF}(q)$ and we can define a homomorphism $\phi:\mathbb{Z}[\alpha]\rightarrow\mathbb{Z}/q\mathbb{Z}$, providing us the representation we aimed for. We denote a prime ideal $s\in O_K$ as \textit{good} if $s$ does not divide the index $[O_K:\mathbb{Z}[\gamma]]$, otherwise \textit{bad}.

The actual factor base $\mathcal{B}$ for which we want to find the logarithms consists of two parts: $\mathcal{B}_\mathbb{Q}$ the set of rational primes and $\mathcal{B}_K$ the set of good prime ideals in $O_K$, where all of these elements are bounded in size, i.e. $p,\Vert p'\Vert_1\leq B,\ \forall p\in\mathcal{B}_\mathbb{Q},p'\in\mathcal{B}_K$. Note that $B$ is the same bound for our base $\alpha$ (from our primary problem definition: Find $D$ for $E=\alpha^D\mod q$). Also, let $\mathcal{B}'\subset\mathcal{B}_\mathbb{Q}$ denote the set of prime factors of $\alpha$.

@@

From here we can derive factorizations for primes in $\mathcal{B}_K$ with the fact (not proven here) that there is an unique ideal $\hat{s}\in\mathcal{B}_K$ lying over $s$ and dividing $c+d\gamma$. Let $v_{\hat{S}}(c,d) = v_s(c,d)$ for this ideal and be zero for other ideals in $\mathcal{B}_K$ of norm $s$. Then we have the following
\begin{eqnarray}
\label{eqnSieve1}
c+dm = \prod_{s\in\mathcal{B}_\mathbb{Q}}s^{w_s(c,d)}
\end{eqnarray}
and
\begin{eqnarray}
\label{eqnSieve2}
c+d\gamma = \prod_{\hat{s}\in\mathcal{B}_K}\hat{s}^{v_{\hat{s}}(c,d)}
\text{ .}
\end{eqnarray}
Repeat the steps described so far to find more pairs $(c,d)$ to build equations \ref{eqnSieve1} and \ref{eqnSieve2} until we have at least $|\mathcal{B}|$ of these equations. For each equation, build a matrix, where its rows are the $w_s$'s and $v_{\hat{s}}$'s, we call this set $\mathcal{A}$. We modify each $A\in\mathcal{A}$ by extracting only those columns where elements are in $\mathcal{B}-\mathcal{B}'$. This means, we are only interested in solutions for primes $s\in\mathcal{B}'$. $\mathcal{A}$ is now the input for the next step.

\subsubsection{Linear Algebra}
For each $A \in \mathcal{A},\ A = \mathbb{Z}^{S\times T},\ S > T$, where each row has at most $X$ non-zero entries, find a linear relation over $\mathbb{Q}$ for the rows of $A$. This is accomplished in four steps, shown for the case $S=T+1$. This procedure was developed by Carl Pomerance (*1944) which was not separately published, instead only for helping out in \citep{gordon1993discrete}.
\paragraph{Step 1}Attempt to compute the rank $r = \mathrm{rk}(A)$. For that, choose a random prime $q_0\leq XT\log T$. Later, we will check if that choice was feasible. If not, come back here and sample another one. Then, use Gaussian elimination mod $q_0$ to find the rank $r_0$ of $A\mod q_0$. Rearrange the rows so that the first $r_0$ rows are linearly independent mod $q_0$. Now the rearranged matrix is
\begin{equation}
A' =
\begin{pmatrix}
v_1\\v_2\\\vdots\\v_{T\!+\!1}
\end{pmatrix}
=
\begin{pmatrix}
a'_{11} & a'_{12} & \cdots & a'_{1T} \\
a'_{21} & a'_{22} & \cdots & a'_{2T} \\
\vdots & \vdots & \ddots & \vdots \\
a'_{T\!+\!11} & a'_{T\!+\!12} & \cdots & a'_{T\!+\!1T} \\
\end{pmatrix}
\text{ .}
\end{equation}
The result of the Gaussian elimination provide a submatrix $\hat{A} = \mathbb{Z}^{r_0\times r_0}$ of the first $r_0$ rows and columns that is nonsingular mod $q_0$,
\begin{equation}
\hat{A} =
\begin{pmatrix}
v_1\\v_2\\\vdots\\v_{r_0}
\end{pmatrix}
=
\begin{pmatrix}
a'_{11} & a'_{12} & \cdots & a'_{1r_0} \\
a'_{21} & a'_{22} & \cdots & a'_{2r_0} \\
\vdots & \vdots & \ddots & \vdots \\
a'_{r_01} & a'_{r_02} & \cdots & a'_{r_0r_0} \\
\end{pmatrix}
\text{ .}
\end{equation}
\paragraph{Step 2} Attempt to express $v_{r_0+1}$ as a linear combination of $v_1,...,v_{r_0}\mod q$ for each $q\leq XT\log T$. For some $q$ it will be successful, for others not. Denote the product of all successful primes as $\boldsymbol{P}$ and the product of all unsuccessful primes as $\boldsymbol{P'}$. If $\boldsymbol{P'} > XT\log T$, then repeat Step 1 (sample a new random prime number). Otherwise, continue with the next step.
\paragraph{Step 3} Attempt to find the determinant $\mathrm{det}(\hat{A})$ of $\hat{A}$. For each prime $q|\boldsymbol{P}$, use the Wieldemann's probabilistic determinant algorithm \citep{wiedemann1986solving} to compute an integer $D_q\in\{0,...,q-1\}$, which is the determinant of $A\mod q$ with probability of at least $1-(XT)^{-2}$, i.e.
\begin{equation}
\mathbb{P}(D_q = \det(A\mod q)\mid A,q) \geq 1-(XT)^{-2}
\text{ .}
\end{equation}
For each $D_q$ we can write an congruence
\begin{equation}
D_0 \equiv D_q \mod q,\quad\forall q|\boldsymbol{P}
\end{equation}
and use the Chinese remainder theorem to compute $D_0$ fulfilling these congruences and being closest to zero. Repeat this step until we have found a value that lies within the bound
\begin{equation}
0 < |D_0| \leq (X^{\frac{1}{2}}T)^T
\text{ .}
\end{equation}

\paragraph{Step 4} In the final step we try to find the actual coefficients $c_1,...,c_{r_0}$ for the linear relation between $v_{r_0+1}$ and the rest of the rows. Again, use the Chinese remainder theorem to find these coefficients closest to zero such that
\begin{equation}
D_0v_{r_0+1}\equiv\sum_{i=1}^{r_0}c_iv_i\mod\boldsymbol{P}
\text{ .}
\end{equation}
In case one of these coefficients $c_i$ exceeds $(X^{\frac{1}{2}}T)^T$ in absolute value, repeat Step 3 to find another $D_0$. Otherwise, we have found the relation
\begin{equation}
D_0v_{r_0+1}=\sum_{i=1}^{r_0}c_iv_i
\text{ .}
\end{equation}
It can also be shown that the expected runtime of this algorithm is
\begin{equation}
\mathcal{O}(X^2T^3\log^3T)
\text{ .}
\end{equation}

@@

After applying the algorithm described in the Linear Algebra step on the matrices $\mathcal{A}$ resulted form the Sieving step, we have some new equations
\begin{equation}
\prod_{(c,d)}(c+dm)^{x(c,d)} = U
\textit{ ,}
\end{equation}
where $x(c,d)$ is the solution for the matrix constructed from $(c,d)$ and $U$ is a unit in $O_K$. After some intermediate steps, we can express the logs of primes $s\in\mathcal{B}$ as $z_s$ (which is a black box here) and state
\begin{equation}
\prod_{s\in\mathcal{B}'}s^{z_s}\equiv 1\mod q
\text{ .}
\end{equation}
Taking the log, we have
\begin{equation}
\sum_{s\in\mathcal{B}'}z_s\log_\alpha s \equiv 0 \mod (q-1)
\text{ .}
\end{equation}
With more than $|\mathcal{B}'|$ such congruences, we can solve these with equation \ref{eqn:assumption} and obtain the logs for all primes $s\in\mathcal{B}'$.
